---
title: "p8105_hw2_ys3298"
author: "Yimeng SHANG"
date: "9/26/2019"
output: github_document
---
# PROBLEM1
```{r}
library(tidyverse)
library(readxl)

trash_wheel = read_excel("../data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1) %>%
  janitor::clean_names() %>%
  filter(dumpster != "NA") %>%
  mutate(sports_balls = as.integer(sports_balls))

##why numeric instead of integer??
mode(trash_wheel$sports_balls) 

```

```{r}

precipi2018 = read_excel("../data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 3, skip = 1) %>% 
  janitor::clean_names() %>%
  filter(total != "NA") %>%
  filter(month != "NA") %>%
  mutate(year = "2018")
total2018 = sum(pull(precipi2018, total))

#why sheet=3 actually tab2?
precipi2017 = read_excel("../data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, skip = 1) %>% 
  janitor::clean_names() %>%
  filter(total != "NA") %>%
  filter(month != "NA") %>%
  mutate(year = "2017")

precipi = bind_rows(precipi2017, precipi2018) %>%
  select(year, month, total) %>%
  mutate(month = month.name[month])
total2017 = sum(pull(precipi2017, total))
```

```{r include=FALSE}
year2018 = filter(trash_wheel, year == "2018")
year2017 = filter(trash_wheel, year == "2017")
year2016 = filter(trash_wheel, year == "2016")
year2015 = filter(trash_wheel, year == "2015")
year2014 = filter(trash_wheel, year == "2014")
```

From the data, we can notice that the total precipitations for 2014-2018 are 51.59, 43.21, 39.95, `r total2017` and `r total2018` respectively, which decrease year by year. The median numbers of sports balls in a dumpster in 2014-2018 are `r median(year2014$sports_balls)`, `r median(year2015$sports_balls)`, `r median(year2016$sports_balls)`, `r median(year2017$sports_balls)`, `r median(year2018$sports_balls)`. The mean numbers of plastic bottle in a dumpster in 2014-2018 are `r median(year2014$plastic_bottles)`, `r median(year2015$plastic_bottles)`, `r median(year2016$plastic_bottles)`, `r median(year2017$plastic_bottles)`, `r median(year2018$plastic_bottles)`. The mean numbers of grocery bags in 2014-2018 are `r round(mean(year2014$grocery_bags), 0)`, `r round(mean(year2015$grocery_bags), 0)`, `r round(mean(year2016$grocery_bags), 0)`, `r round(mean(year2017$grocery_bags), 0)`, `r round(mean(year2018$grocery_bags), 0)`. From these data, we can easily noticed that the number of sports balls, plastic bottle and grocery bags firstly increased and then decrease from 2014 to 2018.

# PROBLEM2
```{r}
pols = read_csv("../data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = month.name[as.factor(month)]) %>%
  select(-prez_dem, -prez_gop, -day)

snp = read_csv("../data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("day", "month", "year")) %>%
  mutate(month = month.name[as.factor(month)]) %>%
  select(year, month, everything())

unemploy = read_csv("../data/unemployment.csv")
tidy_unemploy = 
  pivot_longer(
    unemploy,
    Jan : Dec,
    names_to = "month",
    values_to = "unemployment") %>%
  select(year = Year, everything()) %>%
  mutate( year = as.character(year))

snp_pols = inner_join(snp, pols)
join = left_join(snp_pols, tidy_unemploy, by = c("year", "month"))
```




